---
title: "4211 Homework 5"
author: "Matthew DeSantis"
date: "2023-03-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hmcpkg)
library(BSDA)
library(pwr)
library(haven)
library(faraway)
set.seed(488102)
setwd("C:/Users/Hyrul/Desktop/School/6 Semester/Stats for DS 2/Homework/HW5")
```

# 1

## (a)
```{r 1a}
x = c(19,15,76,23,24,66,27,12,25,7,6,16,51,26,39)
hist(x)
abline(v = 50, col = 'red')
```
## (b)
```{r 1b}
teststat = (2/50)*sum(x)
qchisq(0.95, df = 30) #upper
qchisq(0.05, df = 30) #lower
```
At the 0.10 level of confidence, there is sufficient evidence to reject the claim that theta = 50.

\newpage
# 2
```{r}
power2 = function(theta1){
  a = 0.05
  n = 10
  theta0 = 1
  thing = theta0/theta1  
  upper = thing*qchisq(1-a/2, df = 2*n)
  lower = thing*qchisq(a/2, df = 2*n)
  
  power = 1 - (pchisq(upper, df = 20)-pchisq(lower, df = 20))
  
  return(power)
}

xp = seq(from = 0, to = 5, by = 0.05)
plot(xp, power2(xp))
```

\newpage
# 3

## (a)
(sqrt(n/xbar)*(xbar-theta0))^2

## (b)
```{r 3b}
poiswald = function(x, theta0){
  n = length(x)
  u = mean(x)
  return((sqrt(n/u)*(u-theta0))^2)
}
```

## (c)
```{r 3c}
x = c(27,13,21,24,22,14,17,26,14,22,21,24,19,25,15,25,23,16,20,19)
(stat = poiswald(x, 23))
(1 - pchisq(stat, df = 1))*2
```

\newpage
# 4
```{r 4}
salmon <- read.csv("C:/Users/Hyrul/Desktop/School/6 Semester/Stats for DS 2/Homework/HW5/salmon.dat", sep="")
salmon$R = 1/salmon$recruits
salmon$S = 1/salmon$spawners
lm4 = lm(R~S, data = salmon)
summary(lm4)
```
At the 5% confidence level (but also practically any confidence level, since the pvalues are so small), both beta1 and beta0 are positive.

\newpage
# 5

## (a)
```{r 5a}
data(sat, package="faraway")
afit = lm(total~ratio+salary+expend, data = sat)
summary(afit)
```
From the table, only the intercept is a significant term. Ratio has a positive correlation, salary has a negative, and expend has a positive, though none are significant.

## (b)
```{r 5b}
sat$catratio=(cut(sat$ratio,breaks=c(13,16,18,25)))
bfit = lm(total~catratio+salary+expend, data = sat)
summary(bfit)
```
Here, we treat small ratios (13-16) as the base group. Performing the regression, none of the predictors have significant pvalues, and they are all negative.

## (c)
It changes slightly, in that expend is now negative rather than positive, but it doesn't matter much in practice since they are both insignificant models.